Dispersion distance and the matter distribution of the Universe in dispersion space

### 2015-06-04 writeup/dm_space_distortions.tex
many edits.

%\title{The large-scale structure of the Universe in dispersion measure space}
\title{Dispersion Distance and the Matter Distribution of the Universe in Dispersion Space}

@@

    We propose that `standard pings', brief broadband radio impulses, can be
    used to study the three-dimensional clustering of matter in the Universe
    even in the absence of redshift information.  The dispersion of radio waves
    as they travel through the Universe, like redshift, can be used as a
    cosmological distance measure.  Due to inhomogeneities in the electron
    density along the line-of-siight dispersion is an imperfect proxy for
    radial distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering of sources. Fast radio bursts (FRBs)
    are a new class of radio transients that are the prototypical standard
    ping, and due to their high observed dispersion from intervening plasma,
    have been interpreted as originating at cosmological distances.  The rate
    of fast radio bursts has been estimated to be several thousand over the
    whole sky per day, and if cosmological, the sources of these events should
    trace the large-scale structure of the Universe.  We calculate the
    dispersion-space power spectra for a simple model where electrons and FRBs
    are biased tracers of the large-scale structure of the Universe and show
    that the clustering signal may be detectable in a survey with as few as
    $10000$ events.  Upcoming wide-field radio telescopes, like the Canadian
    Hydrogen Intensity Mapping Experiment (CHIME), are expected to detect...
\end{abstract}

@@

However, redshift is not the only measure of radial distance.   Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to estimate radial distance because of the
characteristic $\lambda^2$ dispersion of electromagnetic signals along the
line-of-sight as a function of wavelength.  This technique, routinely done with pulsars in the Galaxy, can be used to define
a \emph{dispersion distance} to cosmological sources.  Intriguingly, the dispersion of fast radio bursts (FRB), a new class
of millisecond duration radio transients first reported by \citet{2007Sci...318..777L}, has been measured with high significance. 
These bursts are the archetype of a standard ping with 
10 reported
\citep{2012MNRAS.425L..71K,2013Sci...341...53T,
2014ApJ...790..101S, 2014ApJ...792...19B, 
2015MNRAS.447..246P,2015ApJ...799L...5R} to date and total rate of bursts
been estimated to be $3.3^{+5}_{-2.5} \times 10^3$
per sky per day \cite{2015arXiv150500834R}. 
Dispersion is characterized by a line-of-site integral of the free electron density known the dispersion measure
(DM).
FRB events are observed to have dispersion measures of order
$\sim\!\!1000 {\rm pc} / {\rm cm}^3$, greatly in excess of
the galactic expectation of order $\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on
galactic latitude)
\cite{2014arXiv1412.4829D}.
They  thus appear to be of extra-galactic origin, with their
dispersion produced either by the
inter-galactic medium (IGM) \citep{2013Sci...341...53T} surrounding galaxies or the
environment of the source
\citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.
If the dispersion is dominated by electrons along the line-of-sight,
then the average dispersion--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I}
and the sources appear to be at
cosmological distances of of order $\sim\!\!{\rm Gpc}$. Such cosmological distances permit
the study of clustering of FRB sources in three dimensions with an expectation that these events trace the
large-scale structure of the Universe on linear scales. We note that while our discussion here is framed in terms of FRBs, other bright radio transients that fluctuate on timescales less than the typical differential lag of $\sim 1s/100 \,{\rm MHz} $ might also be useful as standard pings.

Neither redshift nor dispersion are perfect proxies for radial distance.
Redshifts are systematically biased by the peculiar velocities of the object
relative to the Hubble flow. This leads to additional apparent clustering
in redshift
space which was first calculated by \citet{1987MNRAS.227....1K}
using linear theory.
Similarly, distance estimates from dispersion will be biased by
inhomogeneities in the electron density \citep{2014ApJ...780L..33M},
again leading to additional apparent
clustering of sources in dispersion measure space.  In this letter we calculate
these dispersion-space distortions and consider the detectability of the signal by
upcoming surveys.

@@

We wish to relate the density of a tracer, $f$, measured in dispersion space to its
density in real space. We follow the derivation in \citet{1987MNRAS.227....1K}
of the
redshift-space distortions. Start by noting that the total number of tracers
in a volume element is the same in both spaces:
\begin{equation}
\label{e:density}
n_{fs}(\vec x_s) \ud^3\vec x_s = n_{f}(\vec x) \ud^3\vec x.
\end{equation}
We split the density into a homogeneous part plus perturbations,
\begin{equation}
    \bar{n}_{fs}(\chi_s)\left[ 1 + \delta_{fs}(\vec x_s)\right] \ud^3\vec x_s
    = \bar{n}_{f}(\chi)\left[ 1 + \delta_f(\vec x)\right] \ud^3\vec x.
\end{equation}
Averaged over the sky, $\langle \chi_s \rangle = \chi$ and thus the background
density should
be the same in dispersion space as in real space,
\begin{equation}
\bar{n}_{fs}(\chi) = \bar{n}_f(\chi).
\end{equation}
Therefore,
\begin{align}
\bar{n}_{fs}(\chi_s) 
    &= \bar{n}_{f}(\chi) + (\chi_s - \chi)\frac{\ud \bar{n}_f}{\ud \chi}\\
    &= \bar{n}_{f}(\chi)
       + \frac{\ud \bar{n}_f}{\ud \chi}\int_0^\chi \ud \chi' \delta_e(\hat n
       \chi').
       \label{e:nfs}
\end{align}

@@

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the equation above, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in dispersion space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to Kaiser.  This is because any wave vector electron perturbation
causes dispertion space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

The $\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}$ arise since the
misinterpretation of the radial distance causes the observed tracer density to
be compared to the wrong background density. The $\frac{2}{\chi}$ term is
caused by a misinterpretation of angular distances when the radial distance is
mis-measured.  In both cases analogous terms are in principle present in 
redshift space but are
negligible. Because radial velocities are only sourced by modes with large
radial wave vector, there is near perfect cancellation along the line of sight
and thus is very little net error in radial distance.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike in redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power-spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

There is an advantage to the lack of kernel in the dispersion-space integral term.
Lensing tomography---the use of sources at multiple redshifts to un-mix
spatial scales---is complicated by the kernel. The kernel's shape depends on
the source redshift, re-weighting the line-of-sight integral and
making tomography inexact.  In contrast the integrand in the dispersion-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus neglect the
information from these slices in our sensitivity measurements and ignore
contributions to the integral term from below this distance. The smallest
spatial scale that contributes the our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

For plotting the dispersion-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB} using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$. The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the form from
\citet{1976ApJ...203..297S} with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting dispersion-space
power-spectrum in Figure~\ref{f:terms}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/n_f}
    \caption{
        \label{f:n_f}
        \emph{Top:} Comoving density of sources in toy model, as described in
        text. Normalization is
        such that the survey has a total of $10^4$ FRB events.
        \emph{Bottom:} Resulting coefficient $A(\chi)$ as given in
        Equation~\ref{e:A}.
    }
\end{figure}

@@

Detailed treatment of the above complications is beyond the scope of this
work. The magnitude of the contribution from the progenitor is at present
unknown. To crudely deal with it here we use relatively large
bins widths of $\Delta\chi = 100\,{\rm Mpc}/h$ 
and exclude the $\chi_i = \chi_j$ bin pairs.
Because the local term drops rapidly with radial separations, this effectively 
discards the contribution from the local term which is most sensitive to having
precise radial distances.

We consider the sensitivity of a survey with a total of $10^4$ dispersion
measures from FRB events distributed uniformly over half the sky. This is in rough
approximation to what could be achieved by the Canadian Hydrogen Intensity
Mapping Experiment's now funded FRB instrument over a three year survey.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights in the radial bin average, we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
The result is shown
in Figure~\ref{f:sensitivity}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/sensitivity}
    \caption{
        \label{f:sensitivity}
        Sensitivity of a survey with $10^4$ dispersion measures distributed
        uniformly over half the sky.  Plotted is the cross-correlation
        power spectrum weighted averaged over all pairs of radial bins. Weights
        are chosen to maximize signal to noise at $\ell=100$.
    }
\end{figure}

@@

We have proposed using FRB sources as 3D tracers of the large-scale structure,
using dispersion to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could be used
to study the distribution of free electrons in the Universe on large-scales.
There have been other proposals to use FRB events to study large-scale
structure \citep{2014ApJ...780L..33M,2014MNRAS.443L..11D}, however these
schemes have either been two dimensional or required externally measured redshifts
for the sources. Additionally, there have been proposals to use the dispersion-redshift
relation to probe the background density of electrons and thus the expansion
history of the Universe
\citep{2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}.

We have
shown that a survey detecting $10^4$ fast radio bursts of cosmological
origin (as might be achieved by the CHIME FRB instrument) could detect the
clustering signal.
Looking forward, dispersion-space
clustering could become a powerful probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.

>>>>

%\title{The large-scale structure of the Universe in dispersion measure space}
\title{Dispersion distance and the matter distribution of the Universe in dispersion space}

@@

    We propose that `standard pings', brief broadband radio impulses, can be
    used to study the three-dimensional clustering of matter in the Universe
    even in the absence of redshift information.  The dispersion of radio waves
    as they travel through the intervening plasma can,
    like redshift, be used as a
    cosmological distance measure.  Due to inhomogeneities in the electron
    density along the line-of-sight, dispersion is an imperfect proxy for
    radial distance and we show that this leads to calculable dispersion-space
    distortions in the apparent clustering of sources. Fast radio bursts (FRBs)
    are a new class of radio transients that are the prototypical standard
    ping, and due to their high observed dispersion,
    have been interpreted as originating at cosmological distances.  The rate
    of fast radio bursts has been estimated to be several thousand over the
    whole sky per day, and if cosmological, the sources of these events should
    trace the large-scale structure of the Universe.  We calculate the
    dispersion-space power spectra for a simple model where electrons and FRBs
    are biased tracers of the large-scale structure of the Universe and show
    that the clustering signal could be measured using as few as
    $10000$ events. Such a survey is in-line with what may be achieved with
    upcoming wide-field radio telescopes such as the Canadian
    Hydrogen Intensity Mapping Experiment (CHIME).
\end{abstract}

@@

However, redshift is not the only measure of radial distance.   Standard
candles and standard rulers can be used to estimate distance using brightness
and angular size respectively.  In the same way a \emph{standard ping}, a brief
broadband radio impulse, can be used to estimate radial distance because of the
characteristic $\lambda^2$ dispersion of electromagnetic signals along the
line-of-sight as a function of wavelength.  This technique, routinely done with
pulsars in the Galaxy, can be used to define a \emph{dispersion distance} to
cosmological sources.  Intriguingly, the dispersion of fast radio bursts (FRB),
a new class of millisecond duration radio transients first reported by
\citet{2007Sci...318..777L}, has been measured with high significance.  These
bursts are the archetype of a standard ping with 10 reported to date
\citep{2012MNRAS.425L..71K,2013Sci...341...53T, 2014ApJ...790..101S,
2014ApJ...792...19B, 2015MNRAS.447..246P,2015ApJ...799L...5R}, and with a total
event rate estimated to be $3.3^{+5}_{-2.5} \times 10^3$ per sky per
day \cite{2015arXiv150500834R}.  Dispersion is characterized by a line-of-site
integral of the free electron density known the dispersion measure (DM).  FRB
events are observed to have dispersion measures of order $\sim\!\!1000 {\rm pc}
/ {\rm cm}^3$, greatly in excess of the galactic expectation of order
$\sim\!\!100\,{\rm pc} / {\rm cm}^3$ (depending on galactic latitude)
\cite{2014arXiv1412.4829D}.  They thus appear to be of extra-galactic origin,
with their dispersion produced either by the inter-galactic medium (IGM)
\citep{2013Sci...341...53T} surrounding galaxies along the line-of-sight
or the environment of the
source \citep{2014ApJ...785L..26L,2015arXiv150101341P,2015arXiv150505535C}.  If
the dispersion is dominated by electrons along the line-of-sight, then the
average dispersion--distance relation can be modelled
\citep{2014ApJ...797...71Z,2004MNRAS.348..999I, 2003ApJ...598L..79I} and the
sources appear to be at cosmological distances of of order $\sim\!\!{\rm Gpc}$.
Such cosmological distances permit the study of clustering of FRB sources in
three dimensions with an expectation that these events trace the large-scale
structure of the Universe on linear scales. We note that while our discussion
here is framed in terms of FRBs, other bright radio transients that fluctuate
on timescales less than the typical differential lag of $\sim 1s/100 \,{\rm
MHz} $ might also be useful as standard pings.

Neither redshift nor dispersion are perfect proxies for radial distance.
Redshifts are systematically biased by the peculiar velocities of objects
relative to the Hubble flow. This leads to additional apparent clustering
in redshift
space which was first calculated by \citet{1987MNRAS.227....1K}
using linear theory.
Similarly, as we show here, distance estimates from dispersion will be biased by
inhomogeneities in the electron density \citep{2014ApJ...780L..33M},
again leading to additional apparent
clustering of sources in dispersion measure space.  In this letter we calculate
these dispersion-space distortions and consider the detectability of the signal by
upcoming surveys.

@@

We wish to relate the density of a tracer, $f$, measured in dispersion space to its
density in real space. We follow the derivation in \citet{1987MNRAS.227....1K}
of the
redshift-space distortions. Start by noting that the total number of tracers
in a volume element is the same in both spaces:
\begin{equation}
n_{fs}(\vec x_s) \ud^3\vec x_s = n_{f}(\vec x) \ud^3\vec x.
\end{equation}
We split the density into a homogeneous part plus perturbations,
\begin{equation}
\label{e:density}
    \bar{n}_{fs}(\chi_s)\left[ 1 + \delta_{fs}(\vec x_s)\right] \ud^3\vec x_s
    = \bar{n}_{f}(\chi)\left[ 1 + \delta_f(\vec x)\right] \ud^3\vec x.
\end{equation}
Averaged over the sky, $\langle \chi_s \rangle = \chi$ and thus the background
density should
be the same in dispersion space as in real space,
\begin{equation}
\bar{n}_{fs}(\chi) = \bar{n}_f(\chi).
\end{equation}
Therefore,
\begin{align}
\bar{n}_{fs}(\chi_s) 
    &= \bar{n}_{f}(\chi) + (\chi_s - \chi)\frac{\ud \bar{n}_f}{\ud \chi}\\
    &= \bar{n}_{f}(\chi)
       + \frac{\ud \bar{n}_f}{\ud \chi}\int_0^\chi \ud \chi' \delta_e(\hat n
       \chi').
       \label{e:nfs}
\end{align}

@@

Substituting Equations~\ref{e:nfs}~and~\ref{e:jac} into Equation~\ref{e:density},
we obtain
\begin{equation}
\label{e:delta_s}
    \delta_{fs} = \delta_f - \delta_e
    - \left(\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}
    + \frac{2}{\chi} \right)
        \int_0^\chi \ud \chi' \delta_e(\hat n \chi').
\end{equation}
In the equation above, the $-\delta_e$ term is most analogous to the Kaiser
redshift-space distortions.  It is a dilution of tracers in dispersion space due to an
excess of electrons between the tracers. However we note that this term is isotropic
in contrast to the Kaiser term.  This is because any wave vector electron perturbation
causes dispertion space distortions, whereas the radial velocities that cause
redshift-space distortions are only sourced by perturbations with radial wave
vectors.

The $\frac{1}{\bar{n}_f}\frac{\ud \bar{n}_f}{\ud \chi}$ term arises because the
misinterpretation of the radial distance causes the observed tracer density to
be compared to the wrong background density. The $\frac{2}{\chi}$ term is
caused by a misinterpretation of angular distances when the radial distance is
mis-measured.  In both cases analogous terms are in principle present in 
redshift space but are
negligible. Because radial velocities are only sourced by modes with large
radial wave vector, there is near perfect cancellation along the line of sight
and thus is very little net error in radial distance.

@@

Large-scale structure is usually studied through its two-point statistics, most
commonly the power spectrum, $P(k)$.
Unlike the redshift-space distortions, Equation~\ref{e:delta_s} does not have a
simple form in harmonic space. The equation's third term couples harmonic modes
and thus the two-point statistics cannot be phrased as a simple power spectrum.
We will instead use $C^{ss}_\ell(\chi,\chi')$, which is
the cross-correlation angular power spectrum of the dispersion-space over density,
on shells at $\chi$ and $\chi'$:
\begin{align}
    &C^{ss}_\ell(\chi, \chi') =
    \int\ud\Omega\ud\Omega' Y_{\ell m}(\hat n) Y^*_{\ell m}(\hat n')
    \langle \delta_{fs}(\hat n \chi) \delta_{fs}(\hat n' \chi')
        \rangle.
\end{align}
The first two terms in Equation~\ref{e:delta_s} are stationary. For stationary,
isotropic, tracers $x$ and $y$ we have 
$\langle \delta_x(\vec k, \chi) \delta_y(\vec k', \chi) \rangle = (2\pi)^3
\delta^3(\vec k - \vec k') P_{xy}(k, \chi)$.  If for a moment we ignore time
dependence, the angular cross-power spectrum of such tracers is
\begin{align}
    %\delta_{\ell\ell'}\delta_{mm'}C^{xy}_\ell(\chi, \chi')
    %&= \int\ud\Omega\ud\Omega'Y_{\ell m}(\hat n) Y_{\ell' m'}(\hat n')
    %    \langle \delta_x(\hat n \chi) \delta_y(\hat n' \chi') \rangle,
    %    \\
    C^{xy}_\ell(\chi,\chi')
    &= \frac{2}{\pi}
\int_0^\infty\ud k k^2 j_\ell(k\chi) j_{\ell}(k\chi')P_{xy}(k).
\end{align}
In reality the power spectrum evolves on the order of a Hubble time.
The angular cross correlations will be very small unless $\chi$ and $\chi'$ are
within a few correlation lengths of one another, roughly a hundred
megaparsecs.  The evolution of the power spectrum is negligible over these time
differences which leads to a straight-forward way to include the evolution of
the perturbations:
\begin{align}
C^{xy}_\ell(\chi,\chi') 
    \approx \frac{2}{\pi}
    \int_0^\infty\ud k k^2
    j_\ell(k\chi) j_{\ell}(k\chi')
    P_{xy}(k,(\chi + \chi')/2),\nonumber &\\
    |\chi - \chi'| \ll 1/aH.&
\end{align}

@@

There is an advantage to the lack of kernel in the dispersion-space integral term.
Lensing tomography---the use of sources at multiple redshifts to un-mix
spatial scales---is complicated by the kernel. The kernel's shape depends on
the source redshift, re-weighting the line-of-sight integral and
making tomography inexact.  In contrast, the integrand in the dispersion-space integral
term is completely independent of the limit of integration, meaning
contributions from different parts of the line of sight (and thus different
spatial scales) can be separated exactly. A detailed study of dispersion space
tomography is beyond the scope of this work.  To avoid our results being
sensitive to the hard-to-model small scales, we assume that the contribution to the
integral term from radial distances $\chi'' < 500\,{\rm Mpc}/h$ are well
measured by the correlations with these radial slices. We thus neglect the
information from these slices in our sensitivity measurements and ignore
contributions to the integral term from below this distance. The smallest
spatial scale that contributes the our plotted angular power spectra out to
$\ell=1000$ is thus $k=2\,h/{\rm Mpc}$.

For plotting the dispersion-space power spectra we use a toy model for the electron and
FRB clustering that assumes both are biased tracers of the dark matter:
\emph{i.e.}~$P_{xy}(k)=b_x b_y P(k)$.  We
calculate the time dependant dark-matter power spectrum using {\tt CAMB} using
the integrated {\tt HALOFIT} to approximate the non-linear evolution. We assume
that the electron bias is $b_e=1$, which should be true on large scales due to
the approximate conservation of free electrons. The bias of FRB sources is
unconstrained by data and we choose it to be $b_f=1.3$. The coefficient
$A(\chi)$ depends on the density of detected FRB sources, $\bar n_f(\chi)$.
We use a simple model where all events above a fixed flux are detected,
the intrinsic luminosity function has the form from
\citet{1976ApJ...203..297S} with index $-1$, and events at the cut off of the
luminosity function and at
a radial distance of $\chi=2350\,{\rm Mpc}/h$ are at the flux detection threshold.
This source density is shown in Figure~\ref{f:n_f} with the resulting dispersion-space
power spectrum in Figure~\ref{f:terms}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/n_f}
    \caption{
        \label{f:n_f}
        \emph{Top:} Comoving density of sources in toy model, as described in
        text. Normalization is
        such that the survey has a total of $10000$ FRB events.
        \emph{Bottom:} Resulting coefficient $A(\chi)$ as given in
        Equation~\ref{e:A}.
    }
\end{figure}

@@

Detailed treatment of the above complications is beyond the scope of this
work. The magnitude of the contribution from the progenitor is at present
unknown. To crudely deal with it here we use relatively large
bins widths of $\Delta\chi = 100\,{\rm Mpc}/h$ 
and exclude the $\chi = \chi'$ bin pairs.
Because the local term drops rapidly with radial separations, this effectively 
discards the contribution from the local term which is most sensitive to having
precise radial distances.

We consider the sensitivity of a survey with a total of $10000$ dispersion
measures from FRB events distributed uniformly over half the sky. This is in rough
approximation to what might be achieved by the Canadian Hydrogen Intensity
Mapping Experiment (CHIME) telescope over a three year survey.
To give an idea of the overall sensitivity of the survey, we collapse
$C^{ss}_{\ell,ij}$ to a single function of $\ell$ by taking a weighted average
of all pairs of radial bins over the
range $500\,{\rm Mpc}/h$ to $3500\,{\rm Mpc}/h$.
For weights we use the signal over noise-squared
at $\ell=100$, which maximizes the signal to noise at that $\ell$ and
is near optimal at other multipoles.
The result is shown
in Figure~\ref{f:sensitivity}.

\begin{figure}
    \includegraphics[scale=0.42]{figures/sensitivity}
    \caption{
        \label{f:sensitivity}
        Sensitivity of a survey with $10000$ dispersion measures distributed
        uniformly over half the sky.  Plotted is the cross-correlation
        power spectrum weighted averaged over all pairs of radial bins. Weights
        are chosen to maximize signal to noise at $\ell=100$.
    }
\end{figure}

@@

We have proposed using FRB sources as 3D tracers of the large-scale structure,
using dispersion to estimate radial distance. It can be seen in
Figure~\ref{f:terms} that the signal is dominated by inhomogeneities in the
electron density along the line of sight inducing mis-estimates of radial
distance and thus apparent clustering. These dispersion-space distortions could be used
to study the distribution of free electrons in the Universe on large-scales and
redshifts less than $\sim\!\!1$.
There have been other proposals to use FRB events to study large-scale
structure \citep{2014ApJ...780L..33M,2014MNRAS.443L..11D}, however these
schemes have either been two dimensional or required externally measured redshifts
for the sources. Additionally, there have been proposals to use the dispersion-redshift
relation to probe the background density of electrons and thus the expansion
history of the Universe
\citep{2014ApJ...783L..35D,2014PhRvD..89j7303Z,2014ApJ...788..189G}.

We have
shown that a survey detecting $10000$ fast radio bursts of cosmological
origin (as might be achieved by the CHIME FRB instrument) could detect the
clustering signal.
Looking forward, dispersion-space
clustering could become a powerful probe of large-scale structure should
observational factors turn out to be favourable compared to spectroscopic
surveys. The baryon acoustic oscillation feature can be clearly seen in
Figure~\ref{f:terms} in the cross term and both
local-term curves. It may also be possible to extract the BAO feature
from the integral
term using tomography.
This however requires that the contribution to dispersion from
the source's environment and host halo be small compared to the
contribution from the IGM.
